Regression Cube Analysis of Cohort Study Data

### 2015-03-29 draft.tex
revision using Sylvias notes

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (left) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $Age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (right) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%% TODO: Cite another paper here!
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover-modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation the system design section later on.
%%
As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, the dynamic variables and the dataset features, many different assumptions can be expressed.
%%
To support \emph{hypothesis-free} analysis, we provide a default formula:

@@

\paragraph{Cube Comparison.}
%%
Regression Cubes can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$)
%%
% \paragraph{Formula Types.}
% %%
% Input formulas reflect \emph{hypotheses} about the data.
% %%
% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
% %%
% Here we show just a few examples on how formulas can serve different purposes.
%
% $Z \sim X + Y$ is the default formula.
% %%
% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
% %%
% Each slice represents a different target feature.
% %%
% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

@@

\section{Summary and Conclusion}
%%
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Model expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.

>>>>

\subsection{Abstracting Regression Results Using $R^2$}
\begin{figure}[htb]
 \centering
 \includegraphics[width=3.5in]{figures/cube}
 \caption{
 (a) Overview visualization using a mosaic plot for the formula $Z \sim X + Y$, where $Z$ assumes the feature $Age$.
 %%
 The $R^2$ values extracted from the regression formulas depict the quality-of-fit and are mapped to color saturation (a saturated color shows a strong correlation).
 %%
 (b) Now, $Z$ is set to all features $n$ and yields $n$ mosaic plot visualizations.
 %%
 These represent the slices in our cube.
 %%
 The $R^2$ values of each slice voxel is mapped on opacity in the 3D-view later on, reducing the occlusion of other values.
 }
  \label{fig:Cube}
\end{figure}
%%
The goal of an overview visualization is to provide a comprehensive view on the data (raw or using descriptive metrics \cite{Bertini}), which is easy to understand.
%% TODO: Cite another paper here!
As described in our previous work \cite{Klemm2014VIS}, correlation values scaled between 0 (no correlation) and 1 (perfect correlation) can be encoded with color in a mosaic plot.
%%
Regression models are more complex, having many associated describing metrics.
%%
For the \emph{Regression Cube} analysis we are interested in the quality-of-fit of the resulting model, which allows to infer about the predictive quality of the independent features included in the model.
%%
As described in Section~\ref{sec:RegressionAnalysis}, the $R^2$ value is the metric allowing for this kind of assessment.
%%
\paragraph{2D (Slice) View.}
%%
Since $R^2$ is scaled between [0,1], it allows for comparison between regression models.
%%
We can apply the same mosaic plot mapping by translating the $R^2$ values to color saturation (Fig.~\ref{fig:Cube} a).
%%
This describes a 2D regression square for dynamic variables $X$ and $Y$ (e.g., $Age \sim X + Y$).
%%

@@

\textbf{Details on Demand} [\textbf{4.DD}]. Precise information about the individual regression models (coefficients, associated confidence intervals and p-values) can be retrieved based on the data point representatives (e.g. in a hover-modal on a currently selected data point).
\\\\
We use the squared bracket abbreviation for each step to denote the affiliation in the system design section later on.
%%
As seen in Fig~\ref{fig:Workflow}, the workflow is highly iterative.
%%
Observations in the 2D mosaic plot or simply the CFS-based features can trigger new analyses by adjusting the underlying regression formulas.
%%
This can be carried out either to refine the current formula based on observations, or creating a new regression cube for a difference view.
%%
\paragraph{Hypothesis-Free and Hypothesis-Based Analysis.}
%%
Input formulas reflect \emph{hypotheses} about the data.
%%
Using the operators, the dynamic variables and the dataset features, many different assumptions can be expressed.
%%
To support \emph{hypothesis-free} analysis, we provide a default formula:

@@

\paragraph{Cube Comparison.}
%%
Regression Cubes can be compared by creating difference views.
%%
One cube (extracted with one formula) acts as reference.
%%
The absolute difference in $R^2$ values towards the second cube is calculated, yielding a difference cube showing only the differences between the two formulas.
%%
For example, it can be utilized for comparing the influence of a single feature towards the complete result (e.g., $Z \sim X + Y$ and $Z \sim X + Y + Income$).
%%
Figure~\ref{fig:Application} (right) shows a comparison cube with subtracted effects of age ($Z \sim Age$).
%%
% \paragraph{Formula Types.}
% %%
% Input formulas reflect \emph{hypotheses} about the data.
% %%
% Using the operators, dynamic variables and dataset features, many different assumptions can be formulated.
% %%
% Here we show just a few examples on how formulas can serve different purposes.
%
% $Z \sim X + Y$ is the default formula.
% %%
% This cube represents all possible combinations of two independent features towards all features in the data set, since we do not know which feature(s) are of interest.
% %%
% Each slice represents a different target feature.
% %%
% It is therefore suitable for a \emph{hypothesis-free} explorative analysis to give a general impression about relationships in the data set.

@@

\section{Summary and Conclusion}
%%
We presented a technique for knowledge discovery in cohort study data sets with user-defined target features.
%%
Dimension reduction using the target restricts the analysis to most important features.
%%
\emph{Hypothesis-free} analysis employs default regression models.
%%
Modeling expert knowledge using regression formulas allow for \emph{hypothesis-based} investigation.
%%
A 3D \emph{Regression Cube} allows to assess hotspots in the analysis by abstracting regression models using a quality-of-fit measure.
%%
These can then be analyzed further using the 2D mosaic plot for each cube slice.
%%
Details on demand for each model allow for detailed assessment of regression models.
%%
We successfully applied the approach to find correlations in a hepatic steatosis as well as a breast cancer data set.
%%
The method was well received by our clinical partners, triggering detailed investigations of the findings.
